---
title: "MBEN - LIANA"
author: "Hanna Schumacher"
output:
  bookdown::html_document2:
    toc: true
    fig_caption: true
    template: flatly
paramsPlot:
  date: !r Sys.Date()
---

# LIANA {.tabset .tabset-fade}

This notebook leads through the ligand-receptor analysis with the tool LIANA. The file *MB_all_t* is used as input.  

The steps are:  

1. Read in data and define filepaths. Create an analysis object that stores this information.
2. Run liana_wrap() on the data and liana_aggregate() on the result. 
3. Run liana_dotplot with the top 25 results according to the aggregate rank.
4. Run liana heat_freq() for aggregate_rank <= 0.01 .


## Libraries
```{r message=FALSE, warning=FALSE}

if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager", version= "1.30.19")

if (!requireNamespace("remotes", quietly = TRUE))
    install.packages("remotes")

BiocManager::install("OmnipathR")
remotes::install_github('saezlab/liana', ref = "0.1.6")

install.packages("../seqDataAnalysisSc", repos = NULL, type="source")

sapply(c(
  "magrittr", "yaml", "functional", "zeallot", "renv",
  "OmnipathR", "liana", "SingleCellExperiment", "Seurat"
  ), seqDataAnalysisSc::loadPackage)


renv::restore()
```

## Paths
```{r}
#' @params 
analysisversion <- "/v00"
host <- "human"
seqType <- "sn"
rel_path_to_proj <- "../projects/MBEN/"
```

```{r}
analysis_params <- yaml::read_yaml(paste0(rel_path_to_proj, "analysis_params_paths.yaml"))
dataset_names <- names(analysis_params$other)[c(1)]
datasetpath_part2 <- paste(analysisversion, host, seqType, sep = "/")
datapath <- analysis_params$default$paths$storagepath_local %>% paste(collapse = "/") %>% paste0(datasetpath_part2)
datasetpath <- analysis_params$default$paths$analysispath_local %>% paste(collapse = "/") %>% paste0(datasetpath_part2)

rm(datasetpath_part2)
```

## Init Analysis Obj.
```{r message=FALSE, warning=FALSE}
# Read Data
datasets <- seqDataAnalysisSc::read_data(dataset_names, datapath, analysis_params, readRDS)
# To SCE
sces <- sapply(datasets, Seurat::as.SingleCellExperiment)
# Init analysis object
analysis <- seqDataAnalysisSc::create_analysis_obj(sces, datasetpath)

rm("analysisversion", "host", "seqType", "rel_path_to_proj", "analysis_params", "dataset_names", "datapath", "datasetpath", "datasets", "sces")
```


Access example: 
`r analysis[["all_t"]]$config$resultpath`

`r table(sces$all_t@colData$PID)`

## Run Liana
```{r}
#' @function run_liana
#' @description Runs liana_wrap & liana_aggregate, if files not already exist, else reads in files.
run_liana <- function(datasetname, analysis){
  dir <- analysis[[datasetname]]$analysis_params$paths$lianadir %>% paste(collapse = "/")
  sce <- analysis[[datasetname]]$data
  res <- list()
  sources <- c("natmi", "connectome", "logfc", "sca", "cellphonedb")
  
  res_dir <- paste0(analysis[[datasetname]]$config$resultpath, "/seurat_clusters/")
  dir.create(res_dir)
  
  filesExist <- sapply(sources, function(sourcename){
    file.exists(paste0(res_dir, sourcename, ".csv")) 
  })
  filesExist <- sum(filesExist) == length(sources)
  if(!filesExist){
    SingleCellExperiment::colLabels(sce) <- sce@colData$seurat_clusters
    res <- sce %>% liana_wrap()
    sapply(names(res), function(sourcename){
      write.csv(res[[sourcename]], paste0(res_dir, sourcename, ".csv"))
    })
    print("Tables are saved.")
  }else{
    res <- sapply(sources, function(sourcename){
      read.csv(paste0(res_dir, sourcename, ".csv"))
    })
    names(res) <- sources
    res <- sapply(res, tibble::as_tibble)
    print("Tables are read in.")
  }
  
  # Aggregate 
  agg <- tibble::tibble()
  agg_dir <- paste0(analysis[[datasetname]]$config$resultpath, "/aggregated/")
  agg_file <- paste0(agg_dir, "aggregation.csv")
  dir.create(agg_dir)

  if(!file.exists(agg_file)){
    agg <- res %>% liana_aggregate()
    write.csv(agg, agg_file)
    agg
  }else{
   agg <- read.csv(agg_file) %>% tibble::as_tibble()
  }
  analysis[[datasetname]]$liana$res <- res
  analysis[[datasetname]]$liana$agg <- agg
  analysis
}
analysis <- run_liana("all_t", analysis)

```

## Dotplot & Heatmap
```{r}

plot_liana_results <- function(datasetname, analysis){
  agg <- analysis[[datasetname]]$liana$agg
  pdf(file = paste0(analysis[[datasetname]]$config$figpath, "/dotplot.pdf"),  width = 12, height = 10)
    clustervalues <- levels(droplevels(as.factor(agg$source)))
    agg$source <- factor(agg$source)
    agg$target <- factor(agg$target)
    agg %>%
        dplyr::top_n(25, dplyr::desc(aggregate_rank)) %>%
        liana_dotplot(source_groups = c(), target_groups = c(clustervalues)) %>% plot
  dev.off()
  print("Dotplot is saved.")
  
  pdf(paste0(analysis[[datasetname]]$config$figpath, "/heatmap.pdf"),  width = 7, height = 6) 
    agg %>% 
      dplyr::filter(aggregate_rank <= 0.01) %>% 
      heat_freq %>% plot
  dev.off()
  print("Heatmap is saved.")
}

sapply(names(analysis), function(datasetname) plot_liana_results(datasetname, analysis))

```

# Seurat -> AnnData 
Conversion for Decoupler analysis.

```{r eval=FALSE, include=FALSE}
#' @function convertFromSeuratToH5ad
convertFromSeuratToH5ad = function (seuratData, datasetpath){
  SeuratDisk::SaveH5Seurat(seuratData, filename = paste0(datasetpath, ".h5Seurat"))
  SeuratDisk::Convert(paste0(datasetpath, ".h5Seurat"), dest = "h5ad")
}
# for each dataset
sapply(names(datasets), function(dataset_name){
  path <-  paste(datapath, dataset_name, "data", dataset_name, sep = "/")
  convertFromSeuratToH5ad(datasets[[dataset_name]], path)
})
```
